DATA 624 Homework 6

### 2020-04-05 DATA624/HW7.Rmd
More work on HW 7

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, 
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(kableExtra)
library(glmnet)
library(pls)
library(corrplot)
library(tidyr)
library(RColorBrewer )
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
```

@@

the `fingerprints` predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.  Filter out the predictors that have low frequencieis using the `nearZeroVar` function from the `caret` package.  How many are left for modeling?

**`r length(nearZeroVar(fingerprints))`**

@@

@@

@@

@@

@@

```{r}

en_predictions <- en_model %>% predict(x_test)
# Model performance metrics
en_results <- data.frame(Model = "Elastic Net Regression",
                         RMSE = caret::RMSE(en_predictions, test_df$y),
                         Rsquared = caret::R2(en_predictions, test_df$y))
en_results %>%
  kable() %>%
  kable_styling()
```

@@

```{r}
varImp(pls_model) %>%
  plot(, top = 20, main = "Important Variables (Cliped at 20)")
```

```{r}
pls_importance <- varImp(pls_model)$importance %>% 
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  arrange(desc(Overall)) %>%
  mutate(importance = row_number())

pls_importance %>%
  mutate(Variable = gsub("[0-9]+", "", Variable)) %>%
  group_by(Variable) %>%
  tally() %>%
  arrange(desc(n)) %>%
  kable() %>%
  kable_styling()
```

**The process predictors dominate the list.**

@@

```{r}
df %>%
  select_at(vars(Yield, pls_importance$Variable)) %>%
  cor() %>%
  corrplot(type = "lower", order = "hclust", tl.col = "black", tl.srt = 45)
  #corrplot(., type = "upper", order = "hclust",
  #       col = brewer.pal(n = 8, name = "RdBu"))
```
  merge(pls_importance) %>%
  ggplot(aes(value, Yield, color = Variable)) +
  theme(legend.position = "none") +
  geom_point() %>%
  facet_wrap(~ Variable)
```

>>>>

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(kableExtra)
library(glmnet)
library(pls)
library(corrplot)
library(tidyr)
library(RColorBrewer)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                     x, perl = TRUE)
  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr("\\]", x)[1] - 1
    # Two cases: start or newline
    re <- paste0("^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub(re, rep,  x)
    re <- paste0("\\\n.{", align_index, "}\\]")
    rep <- paste0("\n", comment)
    x <- gsub(re, rep,  x)
  }

  default_output_hook(x, options)

})

knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))

```

@@

the `fingerprints` predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.  Filter out the predictors that have low frequencies using the `nearZeroVar` function from the `caret` package.  How many are left for modeling?

```{r, null_prefix=TRUE}
fingerprints %>%
  nearZeroVar() %>%
  length()
```

**There are `r length(nearZeroVar(fingerprints))` variables left.**

@@

**I will split it using a 80:20 split.  I will center the data.**

@@

**No improvement here.**

@@

**PLS preformed better.**

@@

**PLS has more explanatory power.**

@@

```{r}
en_predictions <- en_model %>% predict(x_test)
# Model performance metrics
en_results <- data.frame(Model = "Elastic Net Regression",
                         RMSE = caret::RMSE(en_predictions, test_df$y),
                         Rsquared = caret::R2(en_predictions, test_df$y))
en_results %>%
  kable() %>%
  kable_styling()
```

**There isn't any improvment over the PLS.**

@@

**I will select the variables with a `varImp` sore greater than or equal to 60 to be the "important" ones.**

```{r}
pls_importance <- varImp(pls_model)$importance %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  filter(Overall >= 50) %>%
  arrange(desc(Overall)) %>%
  mutate(importance = row_number())

varImp(pls_model) %>%
  plot(., top = max(pls_importance$importance), main = "Important Variables")
```

```{r}
pls_importance %>%
  mutate(Variable = gsub("[0-9]+", "", Variable)) %>%
  group_by(Variable) %>%
  tally() %>%
  arrange(desc(n)) %>%
  kable() %>%
  kable_styling()
```

**There are only `r max(pls_importance$importance)` variables and the process predictors dominate the list.**

@@

```{r}
important_vars <- df %>%
  select_at(vars(Yield, pls_importance$Variable))

important_vars_p <- cor.mtest(important_vars)$p

important_vars %>%
  cor() %>%
  corrplot(method = "color", type = "lower", order = "hclust",
           tl.cex = 0.8, tl.col = "black", tl.srt = 45,
           addCoef.col = "black", number.cex = 0.7,
           p.mat = important_vars_p,  sig.level = 0.05, diag = FALSE)
```

**All variables are positively correlated with the Yield.  Manufacuring process 32 is the most important variable.  It is positively correlated the biological variables.  It is also negatively correlated with manufacturing process 13.**

**Further exploration of these relationships could improve the yield.  This would be valuable information for these chemical manufacturers.**
